Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-10-05 latex/robdml.tex
tex
tables

\usepackage{microtype}
\usepackage{xcolor}
\newcommand{\oo}[1]{\textcolor{orange}{#1}}
\include{math_commands.tex}
\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}
\usepackage{enumitem}

@@

% what is deep metric learning

Given a set of data points, a metric gives a distance value between each pair
of them.
%
Deep Metric Learning (DML) aims to learn such a metric between two inputs (\eg,
images) leveraging the representational power of deep neural networks.
%
DML has been extensively studied~\cite{revisiting}, and has a
wide range of applications such
as image retrieval~\cite{imagesim2} and face recognition~\cite{facenet}.

% why defense in deep metric learning

Despite the improvements of this field thanks to the advancement of deep neural
networks, recent research works suggest that DML models are vulnerable to
adversarial attacks, where an imperceptible perturbation could incur unexpected
or covertly manipulated results~\cite{advrank,advorder}.
%
Such vulnerability poses safety, security, and fairness concerns in the
applications of DML.
%
For example, impersonation or recognition evation are possible for DML-based
face-identification system.
%
To counter the attacks (or reducing the vulnerability), it is important to
design defense methods to improve the adversarial robustness of DML models.

@@

%% for problem 1
%Triplet hardness still matters in adversarial training for deep metric
%learning.
%%
%propose hardness manipulation for.
%solution to problem 1
%
%% for problem 2
%ACT and amdsemi have distinct characteristics.
%solution to problem 2
%
%% for problem 3
%update of advtrain acceleration.
%solution to problem 3
%
%% evaluation and conclusion
%experimetal evaluations

% contributions
%In brief, our contribusions include:
In this MLSP project, we plan to make the following contributions:
%
\begin{enumerate}[noitemsep]
	%
	\item {\textit{Hardness Manipulation}} is proposed for the adversarial
		training of triplet-based deep metric learning models, which avoids
		model collapse in the typical min-max adversarial training setting.
		The proposed method achieves higher adversarial robustness compared to
		the state-of-the-art, and is more computationally efficient.
		%
		This method makes min-max training feaisble, and hence the incorporation
		into free adversarial training and boost learning efficiency.
		%
		(initial condition)
		%
	\item \textit{Gradual Adversary} is proposed to dynamically adjust the
		destination hardness less for Hardness Manipulation during the
		adversarial training process in order to balance the metric learning
		and adversarial learning.
		%
		(terminal condition)
		%
	\item Inter-ID structure constraint. ((aa~p) + (pp~a))/2
		(cross-id repelling)
		existing methods only care inter-class separation, but not
		intra-class (inter-id) separation, because we may get a 6-element
		set (a,a~,p,p~,n,n~) from the original triplet.
		%
	\item Revise and incorporate the state-of-the-art adversarial training
		acceleration method (\ie, Free Adversarial Training~\cite{freeat},
		which is designed for classification) into adversarial training of 
		deep metric learning models. It will greatly improve the efficiency
		of adversarial training.
		%
	\item Benchmark of existing metric learning loss functions under
		adversarial training scenario, and analyze their different
		characteristics for future reference.
		%
	\item Explore the possibility of proposing a new metric learning loss
		function oriented for adversarial training, starting from scratch or
		from an existing metric learning loss.
		%
\end{enumerate}

@@

@@

>>>>

\usepackage{microtype}
\usepackage{xcolor}
\newcommand{\oo}[1]{\textcolor{orange}{#1}}
\include{math_commands.tex}
\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}
\usepackage{enumitem}
\usepackage{multirow}
\newcommand{\uline}[1]{\underline{#1}}

@@

% background and insight.

Given a set of data points, a metric gives a distance value between each pair
of them.
%
Deep Metric Learning (DML) aims to learn such a metric between two inputs (\eg,
images) leveraging the representational power of deep neural networks.
%
DML has been extensively studied~\cite{revisiting}, and has a
wide range of applications such
as image retrieval~\cite{imagesim2} and face recognition~\cite{facenet,domainface}.
%
Despite the improvements of this field thanks to the advancement of deep neural
networks, recent research works suggest that DML models are vulnerable to
adversarial attacks, where an imperceptible perturbation could incur unexpected
or covertly manipulated results~\cite{advrank,advorder}.
%
Such vulnerability poses safety, security, and fairness concerns in the
applications of DML.
%
For example, impersonation or recognition evation are possible for DML-based
face-identification system.
%
To counter the attacks (or reducing the vulnerability), it is important to
design defense methods to improve the adversarial robustness of DML models.

@@

% for problem 1
\oo{[1. Hardness Manipulation]}
Triplet hardness still matters in adversarial training for deep metric
learning.

% for problem 2
\oo{[2. Gradual Adversary]}
The expectation of hardness will decrease during the training process,
compared to that from the beginning phase.

\oo{[3. Intra-Class Structure]}
Given a triplet of samples, we may eventually get up to 6 samples avaiable
for adversarial training. There could be some kind of hierarchy.

\oo{[4. Free Adversarial Metric Learining]}
Extension of FAT to adversarial deep metric learning.

% evaluation and conclusion
\oo{[experimetal evaluations]}
To validate the effectiveness of the proposed method, we conduct experiments
on three commonly used dataset, namely CUB-200-2011, Cars-196, and Stanford
Online Product. \oo{the results suggest that}

\begin{figure}
	\oo{[diagram]}
	\caption{\oo{Diagram demonstrating the core idea}}
\end{figure}

% contributions
In brief, our contribusions include:
%
\begin{enumerate}[noitemsep]
	%
	\item {\textit{Hardness Manipulation}} is proposed for the adversarial
		training of triplet-based deep metric learning models, which avoids
		model collapse in the typical min-max adversarial training setting.
		The proposed method achieves higher adversarial robustness compared to
		the state-of-the-art, and is more computationally efficient.
		%
		This method makes min-max training feaisble, and hence the incorporation
		into free adversarial training and boost learning efficiency.
		%
		(initial condition)
		%
	\item \textit{Gradual Adversary} is proposed to dynamically adjust the
		destination hardness less for Hardness Manipulation during the
		adversarial training process in order to balance the metric learning
		and adversarial learning.
		%
		(terminal condition)
		%
	\item Inter-ID structure constraint. ((aa~p) + (pp~a))/2
		(cross-id repelling)
		existing methods only care inter-class separation, but not
		intra-class (inter-id) separation, because we may get a 6-element
		set (a,a~,p,p~,n,n~) from the original triplet.
		%
	\item Revise and incorporate the state-of-the-art adversarial training
		acceleration method (\ie, Free Adversarial Training~\cite{freeat},
		which is designed for classification) into adversarial training of 
		deep metric learning models. It will greatly improve the efficiency
		of adversarial training.
		%
	\item Benchmark of existing metric learning loss functions under
		adversarial training scenario, and analyze their different
		characteristics for future reference.
		%
	\item Explore the possibility of proposing a new metric learning loss
		function oriented for adversarial training, starting from scratch or
		from an existing metric learning loss.
		%
\end{enumerate}

@@

\oo{[Deep metric learning problem setup]}

In Deep Metric Learning (DML)~\cite{revisiting,dmlreality}, an embedding
function $\phi:\mathcal{X}\mapsto \Phi \subseteq \mathbb{R}^D$ is learned to
map data points $X\in\mathcal{X}$ into an embedding space $\Phi$, where the
output of $\phi(\cdot)$ is usually normalized to the real unit hypersphere for
regularization purpose.
%
With a predefined distance function $d(\cdot,\cdot)$, it allows us to measure
the distance between $X_i$ and $X_j$ as
$d_\phi(X_i,X_j)=d(\phi(X_i),\phi(X_j))$.
%
Various loss functions for learning the underlying embedding function have been
proposed~\cite{revisiting,dmlreality}, where the triplet loss~\cite{facenet}
remain to be a strong and widely used baseline that could reach
state-of-the-art performance with appropriate triplet sampling strategy.

\oo{[Quick review of existing defense methods]}

@@

\subsection{Intra-Class Structure}
